Bioverse: Potentially Observable Exoplanet Biosignature Patterns Under the UV Threshold Hypothesis for the Origin of Life

### 2025-03-27 src/tex/ms.tex
fix a typo, add a citation

\section{Introduction}
\label{sec:intro}
\rev{The probability of abiogenesis remains one of the fundamental questions in modern science.
As of now, it has been primarily explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024}.
At the same time, a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This "UV Threshold Hypothesis" states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

\begin{figure}
    \script{inhabited_aafo_nuv_thresh.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate $f_\mathrm{life} = 1$.
        %Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
        For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
        \label{fig:inhabited_aafo_nuv_thresh}
    \end{centering}
\end{figure}
Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for the limiting case of \rev{a probability for life's emergence and persistence} of $f_\mathrm{life} = 1$.
This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs.
For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.

@@

On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes~\citep{Richey-Yowell2023}.
This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher \gls{NUV} flux is favorable or necessary for life.
At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.

@@

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/combined_fluxes.pdf}
    \caption{\rev{Comparison of \gls{NUV} flux evolution in terms of photon number flux and energy flux for K-type, early M-type, and late M-type stars.
    In both regimes, later spectral types exhibit higher maximum values.
    The ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs is slightly larger in the photon number flux regime compared to the energy flux regime, suggesting that our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis are conservative.
    }}
    \label{fig:combined_fluxes}
\end{figure*}
\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution (Figure~\ref{fig:combined_fluxes}).
Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities }\revv{due to the steeper rise of the blackbody SEDs at shorter wavelengths}.

>>>>

\section{Introduction}
\label{sec:intro}
\rev{The probability of abiogenesis remains one of the fundamental questions in modern science.
As of now, it has been primarily explored through statistical arguments~\citep[e.g.,][]{Spiegel2012,Kipping2021,Lingam2024}.
At the same time, a} wide variety of scenarios for the origin of life have been proposed~\citep[e.g.,][]{Baross1985,Brasier2011,Mulkidjanian2012,Fox2013,Deamer2015,Westall2018}.
While we may still be far from conclusively testing these scenarios, new prospects in the search for conditions favorable to life have opened up by thinking of the origin of life as a planetary phenomenon and identifying global-scale environmental properties that might support pathways to life~\citep{Sasselov2020}.
In particular, specific planetary conditions are needed to create stockpiles of initial compounds for prebiotic chemistry; and planetary processes are required to trigger the prebiotic synthesis.
Such planetary conditions can be hypothesized for exoplanets located in the \gls{HZ} of their host star, with persistent liquid water on their surface.
For example, if deep-sea or sedimentary hydrothermalism is required for abiogenesis, then the insulation of an ocean from the planetary crust minerals (e.g., due to high-pressure ices) may reduce or eliminate the chances of life emerging~\citep[e.g.,][]{Baross1985}.
The alternate scenario of a surface locally subject to wet-dry cycles requires a planetary exposure to mid-range \gls{UV} irradiation, as a source of energy and an agent of selection in chemical evolution~\citep[e.g.,][]{Deamer2019}.
This ``UV Threshold Hypothesis'' states that \gls{UV} light in a specific wavelength range played a constructive role in getting life started on Earth~\citep{Ranjan2016,Ranjan2017c,Rimmer2018,Rapf2016}, and it could provide a probabilistic approach to the interpretation of possible future biosignature detections~\citep[e.g.,][]{Catling2018,Walker2018}.

@@

\begin{figure}
    \script{inhabited_aafo_nuv_thresh.py}
    \begin{centering}
        \includegraphics[width=\hsize]{figures/inhabited_aafo_nuv_thresh}
        \caption{Fraction of inhabited planets for different threshold \gls{NUV} fluxes under the UV~Threshold Hypothesis if the abiogenesis rate $f_\mathrm{life} = 1$.
        %Shaded regions denote 90 percent confidence intervals of randomized sample generations, and dashed lines correspond to samples including non-transiting planets.
        For all samples, the fraction of inhabited planets drops sharply with increasing threshold \gls{NUV} flux due to the combined effects of never receiving sufficient \gls{NUV} flux for abiogenesis or receiving it before entering the \gls{HZ}.}
        \label{fig:inhabited_aafo_nuv_thresh}
    \end{centering}
\end{figure}
Figure~\ref{fig:inhabited_aafo_nuv_thresh} shows the fraction of inhabited planets under the UV~Threshold Hypothesis for different threshold \gls{NUV} fluxes and for the limiting case of \rev{a probability for life's emergence and persistence} of $f_\mathrm{life} = 1$.
This fraction decreases sharply with increasing threshold flux, as fewer planets receive sufficient \gls{NUV} flux for abiogenesis.
Another effect responsible for this drop is that some planets receive the required \gls{NUV} flux only before entering the \gls{HZ} -- this is especially likely for M~dwarfs\revv{~\citep{Spinelli2024}}.
For the FGK sample, the fraction of inhabited planets drops at lower threshold fluxes than for the M dwarf sample.

@@

On the other hand, M~dwarfs show a wider distribution of maximum past \gls{NUV} fluxes in their \glspl{HZ}.
While old M~dwarfs can be considered low-UV environments, a significant fraction of them emit high \gls{NUV} fluxes into their \gls{HZ} during their early stages, in particular later subtypes\revv{~\citep{Richey-Yowell2023,Spinelli2024}}.
This will help to test the high \gls{NUV} flux end of the UV Threshold Hypothesis; a higher occurrence of biosignatures here would support the hypothesis that a higher \gls{NUV} flux is favorable or necessary for life.
At the same time, a fraction of host stars in our M~dwarf sample extends it to lower maximum past \gls{NUV} fluxes, enabling tests of the low \gls{NUV} flux end of the hypothesis.
The higher and more variable \gls{NUV} fluxes in M~dwarfs thus increase the likelihood of obtaining strong evidence for or against the UV~Threshold Hypothesis.

@@

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/combined_fluxes.pdf}
    \caption{\rev{Comparison of \gls{NUV} flux evolution in terms of photon number flux and energy flux for K-type, early M-type, and late M-type stars.
    In both regimes, later spectral types exhibit higher maximum values.
    The ratio of the maximum past \gls{NUV} fluxes in K dwarfs and late M dwarfs is slightly larger in the photon number flux regime compared to the energy flux regime, suggesting that our conclusions regarding the impact of host star spectral type on the UV Threshold Hypothesis are conservative.
    }}
    \label{fig:combined_fluxes}
\end{figure*}
\rev{We found that the derived flux densities for all considered spectral types exceed the threshold estimated by \citet{Rimmer2018} during earlier stages of stellar evolution\revv{~\citep[Figure~\ref{fig:combined_fluxes}, also see][]{Spinelli2024}}.
Notably -- as in the case of energy fluxes -- later spectral types exhibit higher maximum past \gls{NUV} number flux densities }\revv{due to the steeper rise of the blackbody SEDs at shorter wavelengths}.
